---
title: "STAT 685 / TCGA Lymphovascular Pheatmap"
author: "Eric Holdaway / eric.holdaway@tamu.edu"
date: "8/3/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This code is producing PHeatmaps for our data based upon Lymphovascular Invasion, currently it is showing the clustering based of of Hierarchical Complete Clustering. Included in this is the Adjusted Rand Indexing values to get an idea of how well the data fits and compare it to a few other techniques.  

**Creates the pheatmap for the following:**  
Complete Case Clustering  
Clustering with missing up to 60% of the values  
Clustering using imputed data from missForest()  

**Also includes the following:**  
Fuzzy C-Means Clustering
Hierarchical Complete Clustering
K-means Clustering  
Kpod Clustering (cluster with missing data)    

### Libraries:
kpodclustr  
fossil  
pheatmap  
ppclust  

```{r Library,include=FALSE}
library(kpodclustr) # kpod()
library(fossil)     # adj.rand.index()
library(pheatmap)   # pheatmap()
library(ppclust)    # fcm()
```

### Data:  

```{r Data}
setwd("C:/Users/erich/OneDrive/Documents/School/STAT 685/Data")
# Had to change the "-" in the PatientsID to "_" so mice would work.
data_base <- read.csv("DataBase_lymph.csv",header=TRUE)
data_clinic <- read.csv("DataClinic_lymph.csv",header=TRUE)
data_imp <- read.csv("Data_Impute_Lymph.csv",header=TRUE)
```

### Seperation of data:  
Here I am setting the number of cluster for the rest of the code, based upon the number of different Lymphovascular Invasion values = 2. Also, I am cleaning up and transforming the data and data frame into a clean working version for the rest of the code.  

```{r Data Cleaning 1}
# Creation of gene and variable array's.
# Cleans out the first row of count data.
data_base <- data_base[,-1]
data_variables <- data_base[,1]
data_genes <- colnames(data_base)

# Creation of data set.
data_values <- data_base[,-1]
data_trans <- data_values

# Setup of variables for complete case analysis.
data_complete <- data_trans
clinic_complete <- data_clinic
variables_complete <- data_variables
genes_complete <- data_genes

# Setup of variables for missForest imputed data.
clinic_imp <- data_clinic
variables_imp <- data_variables
genes_imp <- data_genes

# Variables.
n_clust <- 2
```


### Removal of Genes with all NA's:  
This step is for all the genes which have nothing but NA values.

```{r Data Cleaning 2}
# Variables for first check.
keep <- vector()
perc_na <- vector()

# First check is to find all columns that have nothing but "NA" values and eliminate them.
# Checing by patient results in no patients removed from list. All have at least some data.
for(i in 1:(dim(data_trans)[2]))
{
  val_true <- table(is.na(data_trans[,i]))[2]
  perc_na <- append(perc_na,val_true/dim(data_trans)[1])
}

for(i in 1:(length(perc_na)))
{
  if(is.na(perc_na[i]) != "TRUE")
  {
    keep <- append(keep,i)
  }
}

# All columns with all "NA" values have been removed.
data_keep <- data_trans[,keep]
data_genes <- data_genes[keep]
```

### Removal of Genes with 60% or more NA's removed:  
Here I am removing those genes which have a large percentage of missing data, +60%.  

```{r Data Cleaning 3}
# Variables for second check, percentage level to keep.
perc_na <- vector()

# Second check.
for(i in 1:(dim(data_keep)[2]))
{
  val_true <- table(is.na(data_keep[,i]))[2]
  perc_na <- append(perc_na,val_true/dim(data_keep)[1])
}

keep <- which(perc_na < 0.60)

# Completed data set with only columns with less than 60% NA's remaining.
# Data is missing gene names.
data_keep <- data_keep[,keep]
data_genes <- data_genes[keep]


# Converting array to workable data frame.
data_keep <- as.data.frame(data_keep)

# For loop to convert all columns to numeric.
for(i in 1:dim(data_keep)[2])
{
 data_keep[,i] <- as.numeric(data_keep[,i]) 
}

colnames(data_keep) <- data_genes
rownames(data_keep) <- data_variables

# Creation of two data_keep objects for different testing.
data_keep_df <- data_keep
data_keep_matrix <- as.matrix(data_keep)
```

### Pheatmap for Lymphovascular Invasion:  
Creating Pheatmaps for a visualization of the grouping of data. I have also included a few clustering values along with the heat map so we can get a numeric representation of what is going on a well. Data with no missing values has 3 clustering techniques applied while data with missing values only have one technique applied.  

**Up to 60% missing data:**  

```{r Lymphovascular Invasion Pheatmap}
my_sample_col <- data.frame(data_clinic$lymphovascular_invasion)
row.names(my_sample_col) <- data_clinic$PatientID
colnames(my_sample_col) <- "L.I."

pheatmap(data_keep_matrix, cluster_rows=TRUE, show_rownames=FALSE, show_colnames=FALSE, cluster_cols=FALSE, cutree_rows=n_clust, annotation_row=my_sample_col)

# Rand Indexing to give a numeric comparison.
clust_kpod <- kpod(data_keep_matrix,k=n_clust)
adj.rand.index(clust_kpod$cluster,data_clinic$lymphovascular_invasion)
```

There is no real distinctive grouping and the adjusted rand index values is close to zero telling us that it is similar to random selection.  

**Complete Case Analysis:**  

```{r Data Cleaning Complete 1}
# Converting array to workable data frame.
data_complete <- as.data.frame(data_complete)
rownames(data_complete) <- variables_complete

# For loop to convert all columns to numeric.
for(i in 1:dim(data_complete)[2])
{
 data_complete[,i] <- as.numeric(data_complete[,i]) 
}

data_complete <- data_complete[, colSums(is.na(data_complete)) == 0]
```

```{r L.I. Pheatmap Complete Case}
my_sample_col <- data.frame(clinic_complete$lymphovascular_invasion)
row.names(my_sample_col) <- clinic_complete$PatientID
colnames(my_sample_col) <- "L.I."

pheatmap(data_complete, cluster_rows=TRUE, show_rownames=FALSE, show_colnames=FALSE, cluster_cols=FALSE, cutree_rows=n_clust, annotation_row=my_sample_col)

clust_fuzzy <- fcm(data_complete,centers=n_clust)
adj.rand.index(clust_fuzzy$cluster,data_clinic$lymphovascular_invasion)

hclust_com <- hclust(dist(data_complete), method="complete")
hclust_clust <- cutree(tree=hclust_com, k=n_clust)
adj.rand.index(hclust_clust,data_clinic$lymphovascular_invasion)

clust_kmean <- kmeans(data_complete,centers=n_clust)
adj.rand.index(clust_kmean$cluster,data_clinic$lymphovascular_invasion)
```

Get very similar results as the data with missing values left in and doesn't show us anything useful. One thing to point out is that it appears there is some clear differentiation doing one in the miRNA and if we might get good results using it for clustering. All indexing values are close to zero, again showing that the clustering is similar to random selection.  

**MissForest Imputation (Not Standardized):**  

```{r L.I. Pheatmap missForest 1}
rownames(data_imp) <- data_imp[,1]
data_imp <- data_imp[,-1]

my_sample_col <- data.frame(clinic_imp$lymphovascular_invasion)
row.names(my_sample_col) <- clinic_imp$PatientID
colnames(my_sample_col) <- "L.I."

pheatmap(data_imp, cluster_rows=TRUE, show_rownames=FALSE, show_colnames=FALSE, cluster_cols=FALSE, cutree_rows=n_clust, annotation_row=my_sample_col)

clust_fuzzy <- fcm(data_imp,centers=n_clust)
adj.rand.index(clust_fuzzy$cluster,clinic_imp$lymphovascular_invasion)

hclust_com <- hclust(dist(data_imp), method="complete")
hclust_clust <- cutree(tree=hclust_com, k=n_clust)
adj.rand.index(hclust_clust,clinic_imp$lymphovascular_invasion)

clust_kmean <- kmeans(data_imp,centers=n_clust)
adj.rand.index(clust_kmean$cluster,clinic_imp$lymphovascular_invasion)
```

We have very similar results to the complete case heat map above and in the indexing values. The big difference is that we can no longer see possible groupings in the miRNA.  

**MissForest Imputation (Standardized)**  

```{r L.I. Pheatmap missForest 2}
data_imp <- scale(data_imp)

my_sample_col <- data.frame(clinic_imp$lymphovascular_invasion)
row.names(my_sample_col) <- clinic_imp$PatientID
colnames(my_sample_col) <- "L.I."

pheatmap(data_imp, cluster_rows=TRUE, show_rownames=FALSE, show_colnames=FALSE, cluster_cols=FALSE, cutree_rows=n_clust, annotation_row=my_sample_col)

clust_fuzzy <- fcm(data_imp,centers=n_clust)
adj.rand.index(clust_fuzzy$cluster,clinic_imp$lymphovascular_invasion)

hclust_com <- hclust(dist(data_imp), method="complete")
hclust_clust <- cutree(tree=hclust_com, k=n_clust)
adj.rand.index(hclust_clust,clinic_imp$lymphovascular_invasion)

clust_kmean <- kmeans(data_imp,centers=n_clust)
adj.rand.index(clust_kmean$cluster,clinic_imp$lymphovascular_invasion)
```

We have very similar results as the previous Pheatmaps and the indexing values. Overall there appears to be no distinct grouping in the patients based upon Lymphovascular Invasion.  